
title: "Data Cleaning: Cost and Affordability of a Healthy Diet"
author: "Ella Kulicki"
format:
    html:
        embed-resources: true
---

## Datasets to Import and Clean
1. Cost and Affordability of a Healthy Diet - done
2. Availability (of different nutrients)

# Cost and Affordability of a Healthy Diet

Elements (with unit):
- Prevalence of unaffordability (%)
- Number of people unable to afford a healthy diet (number in millions)


```{r}

library(tidyverse)
library(dplyr)
library(tidyr)


cahd_raw <- read.csv("raw_data/FAOSTAT_cost_and_affordability.csv", stringsAsFactors = FALSE)

# keep only necessary columns
cahd_df <- cahd_raw %>%
    select(Area, Year, Item, Value) %>% 
    rename(
        country = Area,
        year = Year,
        item = Item,
        value = Value
    )

# pivot columns to rows
cahd_df <- cahd_df %>%
    pivot_wider(
        id_cols = c(country, year),  
        names_from = item,           
        values_from = value         
    )

# rename new columns
cahd_df <- cahd_df %>%
    rename(
        prevalence_of_unaffordability = `Prevalence of unaffordability (PUA), percent`,
        num_of_ppl_unable_to_afford_healthy_diet = `Number of people unable to afford a healthy diet (NUA), million`
    )

# convert to numeric

cahd_df <- cahd_df %>%
  mutate(
    year = as.numeric(year),
    prevalence_of_unaffordability = as.numeric(prevalence_of_unaffordability),
    num_of_ppl_unable_to_afford_healthy_diet = as.numeric(num_of_ppl_unable_to_afford_healthy_diet)
  )


write.csv(cahd_df, "clean_data/cost_and_affordability_clean.csv", row.names = FALSE)




```

# Availability (of different nutrients)


```{r}
#| eval: true 

library(tidyverse)
library(dplyr)
library(tidyr)


avail_raw <- read.csv("raw_data/availability_data_raw.csv", stringsAsFactors = FALSE)


# keep only necessary columns
avail_df <- avail_raw %>%
    select(Area, Food.Group, Indicator, Y2017, Y2018, Y2019, Y2020, Y2021, Y2022) %>% 
    rename(
        country = Area,
        food_group = Food.Group,
        indicator = Indicator,
        `2017` = Y2017, 
        `2018` = Y2018, 
        `2019` = Y2019, 
        `2020` = Y2020, 
        `2021` = Y2021, 
        `2022` = Y2022
    )

avail_df <- avail_df %>%
    pivot_longer(
        cols = c("2017", "2018", "2019", "2020", "2021", "2022"),
        names_to = "year",
        values_to = "value" 
    )

# keep just "all food groups"
avail_df <- avail_df %>%
    filter(food_group == "All food groups")

avail_df <- avail_df %>%
    pivot_wider(
        id_cols = c("country", "year"),
        names_from = "indicator",
        values_from = "value"
    )
# rename indicator columns

avail_df <- avail_df %>%
    rename(
        energy_supply = `Energy supply`,
        protein_supply = `Protein supply`,
        fat_supply = `Fat supply`,
        carbohydrate_supply = `Carbohydrate (available) supply`,
        dietary_fibre_supply = `Dietary fibre supply`,
        calcium_supply = `Calcium supply`,
        iron_supply = `Iron supply`,
        zinc_supply = `Zinc supply`,
        magnesium_supply = `Magnesium supply`,
        phosphorus_supply = `Phosphorus supply`,
        potassium_supply = `Potassium supply`,
        thiamin_supply = `Thiamin supply`,
        riboflavin_supply = `Riboflavin supply`,
        vitamin_c_supply = `Vitamin C supply`,
        vitamin_a_supply_retinol_equivalents = `Vitamin A supply (retinol equivalents)`,
        vitamin_a_supply_retinol_activity_equivalents = `Vitamin A supply (retinol activity equivalents)`,
        copper_supply = `Copper supply`,
        docosahexaenoic_acid_supply = `Docosahexaenoic acid (DHA) supply`,
        vitamin_b6_supply = `Vitamin B6 supply`,
        eicosapentaenoic_acid_supply = `Eicosapentaenoic acid (EPA) supply`,
        selenium_supply = `Selenium supply`,
        vitamin_b12_supply = `Vitamin B12 supply`,
        total_saturated_fatty_acids_supply = `Total saturated fatty acids supply`,
        total_monounsaturated_fatty_acids_supply = `Total monounsaturated fatty acids supply`,
        total_polyunsaturated_fatty_acids_supply = `Total polyunsaturated fatty acids supply`
    ) %>%
    mutate(
        year = as.numeric(year)
    )


write.csv(avail_df, "clean_data/availability_clean.csv", row.names = FALSE)


```
