---
title: "Expenditure and Investment to Agriculture Data Cleaning"
author: "Sophie McDowall"
format:
    html:
        embed-resources: true
---

## Data and Plan

Datasets to Import and Clean
To import and clean:
1. Government Expenditure
2. Development Flows to Agriculture
3. Foreign Direct Investment (https://www.fao.org/faostat/en/#data/FDI)
4. Production Data

- [x] Government Expenditure
- [x] Development Flows to Agriculture
- [ ] Foreign Direct Investment
- [ ] Production Data



Process:
- import from raw_data folder <br>
- clean <br>
- reorder into country/year units <br>
- join all datasets <br>
- export to clean_data folder <br>

Output:
Two .csv files uploaded to clean_data folder, including one with the years 2001-2022 and one with only the years 2017-2022

```{r}
#imports
library(tidyverse)
library(dplyr)

```

### Development Flows to Agriculture
https://www.fao.org/faostat/en/#data/EA 

Output definition:
Year
Country: recipient country of money
Total Disb Received: total disbursementes received from all donors in USD, 2022 USD

```{r}
flow_df <- read.csv("raw_data/FAOSTAT_devflows.csv")
flow_df <- subset(flow_df, select = -c(Domain, Donor, Element, Item, Unit, 
Purpose))
flow_df <- flow_df |> rename("total_disb_received" = "Value")
flow_df <- flow_df |> rename("country" = "Recipient.Country")
flow_df <- flow_df |> rename("year" = "Year")

flow_df_short <- flow_df |>
    filter(year %in% c(
        "2017", "2018", "2019", "2020", "2021", "2022"
    ))

```


### Government Expenditure
https://www.fao.org/faostat/en/#data/EA 

Includes both General Government Total Expenditures and General Government Expenditures for Agriculture, Forestry, and Fishing

Output definition:
Year
Country: recipient country of money
Total Disb Received: total disbursements received from all donors in USD, 2022 USD

```{r}

#first for download of general government total expenditures
expend_df1 <- read.csv("raw_data/FAOSTAT_gengovexpend.csv")
expend_df1 <- subset(expend_df1, select = -c(Domain.Code, Domain, Area.Code..M49., Element.Code, Element, Item, Item.Code, Year.Code, Note, Unit)) 
expend_df1 <- expend_df1 |> rename("total_expenditures" = "Value")
expend_df1 <- expend_df1 |> rename("country" = "Area")
expend_df1 <- expend_df1 |> rename("year" = "Year")

expend_df1_short <- expend_df1 |>
    filter(year %in% c(
        "2017", "2018", "2019", "2020", "2021", "2022"
    ))


#second for download of general government agriculture, forestry, fishing expenditures
expend_df2 <- read.csv("raw_data/FAOSTAT_affgovexpend.csv")
expend_df2 <- subset(expend_df2, select = -c(Domain.Code, Domain, Area.Code..M49., Element.Code, Element, Item, Item.Code, Year.Code, Note, Unit)) 
expend_df2 <- expend_df2 |> rename("aff_expenditures" = "Value")
expend_df2 <- expend_df2 |> rename("country" = "Area")
expend_df2 <- expend_df2 |> rename("year" = "Year")

expend_df2_short <- expend_df2 |>
    filter(year %in% c(
        "2017", "2018", "2019", "2020", "2021", "2022"
    ))

#combine the two based on country 
combined_expend <- full_join(expend_df1, expend_df2, by = c("country", "year"))
combined_expend_short <- full_join(expend_df1_short, expend_df2_short, by = c("country", "year"))


```





## Combining to One DF for Output

```{r}
final_finance_df <- flow_df |>
    full_join(combined_expend, by = c("country", "year")) |>
    full_join(combined_expend, by = c("country", "year"))


final_finance_df_short <- flow_df_short |>
    full_join(combined_expend_short, by = c("country", "year")) |>
    full_join(combined_expend, by = c("country", "year"))


write.csv(final_finance_df, "clean_data/finance_data_01to22.csv", row.names = FALSE)
write.csv(final_finance_df_short, "clean_data/finance_data_17to22.csv", row.names = FALSE)



```




















## Earlier code if needed

library(tidyverse)
library(dplyr)

flow_df <- read.csv("raw_data/Development_Assistance_to_Agriculture_E_All_Data/Development_Assistance_to_Agriculture_E_All_Data_NOFLAG.csv")

#subset to years 2001-2022
flow_df1 <- flow_df[, c(3, 6, 8, 10, 12, 13, which(names(flow_df) == "Y2001"):which(names(flow_df) == "Y2022"))]

#filtering for desired attributes
flow_df2 <- flow_df1 |>
    filter(Purpose == "Agriculture, forestry, fishing",) |>
    filter(Donor == "All Donors") |>
    filter(Element == "Value US$, 2022 prices") |>
    filter(Item == "Disbursement")

#removing unnecessary rows and columns, to only include year columns and rows that are one country
flow_df_final <- flow_df2[1:170, ]
flow_df_final <- subset(flow_df_final, select = -c(Purpose, Donor, Element, Item, Unit))




